---
title: "COMP2501 Assignment 1"
output:
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Requirements

**Submission deadline: Feb 26th, 2026 at 23:59.**

**Full mark of Assignment 1: 50.**

For the following questions, please:

1.  Replace all [Input here] places with your information or your answer (for multiple choice).
2.  Complete the code block by adding your own code to fulfill the requirements in each question. Please use the existing code block and do not add your own code block.

Please make sure your Rmd file is a valid Markdown document and can be successfully knitted.

For assignment submission, please knit your final Rmd file into a Word document, and submit both your **Rmd** file and the knitted **Microsoft Word** document file to Moodle. You get 0 score if 1) the Rmd file you submitted cannot be knitted, and 2) you have not submitted a Word document. For each visualization question, please make sure that the generated plot is shown in-place with the question and after the code block.

------------------------------------------------------------------------

## Name and UID

Name: Kan Chi Fu

UID: 3035807902

------------------------------------------------------------------------

### Environmental setup

You need to have the `dslabs`, `dplyr`, and `ggplot2` packages installed. If not yet, please run `install.packages(c("dslabs", "dplyr", "ggplot2"))` in your R environment. If you have installed the `tidyverse` package, `dplyr` is installed by default.

```{r}

# Load the required packages
# install.packages(c("dslabs", "dplyr", "ggplot2"))
library(dslabs)
library(dplyr)
library(ggplot2)

```

### Example question. Print the first 2 records of the `murders` dataset.

```{r}

data("murders")
head(murders, 2)
```

### 1. (1 points) Given matrix `mat`, print the entries of row 2, and columns 2 to 4. Please define a matrix of your choice. You will need to select both row 2 and columns 2 to 4 simultaneously.

```{r}
mat <- matrix(1:100,10,10)
print(mat[2,2:4])
```

### 2. (2 points) Write a function `compute_s_n` that for any given n, computes the $S_n = n*\sqrt{(n+9)}*log_{10}(n)$. Print the $S_n$ with $n = 500$.

```{r}
compute_s_n <- function(n){
  return(n * sqrt(n+9) * log10(n))
}
print(compute_s_n(500))
```

### 3. (2 points) Examine the built-in dataset `Orange`. Which of the following is true?

-   

    a.  Orange is tidy data: it has one observation for each row.

-   

    b.  Orange is not tidy: we need at least one column with a character vector.

-   

    c.  Orange is not tidy: it is a matrix instead of a data frame.

-   

    d.  Orange is tidy data: all small datasets are tidy by definition.

Your answer is: a

### 4. (15 points) Utilize the `storms` dataset included in the `dplyr` package to analyze tropical cyclone activities. The `storms` dataset contains the tracks of tropical storms in the Atlantic Ocean. You will need to clean, aggregate, and visualize this complex dataset.

```{r}

# Load the required packages and dataset
library(dplyr)
library(ggplot2)
data("storms")
```

#### a. (3 points) Filter the dataset to include only storms from the year 1990 onwards. Create a new column called `storm_category` that classifies storms based on their wind speed (in knots) using the Saffir-Simpson scale below. Treat missing values as "Unknown". Tropical Depression: \< 34 knots. Tropical Storm: 34–63 knots. Hurricane: larger or equal to 64 knots.

```{r}
# 1. Filters for 1990 onwards
storms <- filter(storms, year >= 1990)
# 2. New column "storm_category"
storms <- storms |> mutate(
  storm_category = case_when(
    is.na(wind) ~ "Unknown",
    wind < 34 ~ "Tropical Depression",
    between(wind, 34, 63) ~ "Tropical Storm",
    wind > 63 ~ "Hurricane"
  )
)
print(storms |> group_by(storm_category) |> count())
```

#### b. (4 points) For each unique storm (identified by both name and year), calculate the maximum wind speed and the total duration of the storm (count the number of observations). Identify the top 5 storms with the longest duration that reached "Hurricane" status. Print their names, years, max wind speeds, and duration.

```{r}
# Summarize max speed and duraion
uni_storms <- storms |> group_by(name,year) |> summarise(
  is_hurricane = "hurricane" %in% status,
  max_wind_speed = max(wind),
  duration = n(),
  .groups = "keep"
)
# Get top 5 hurricane
top_hurr <- uni_storms |> filter(is_hurricane) |> arrange(desc(duration)) |> head(5) |>
  select(name,year,max_wind_speed,duration)
print(top_hurr)
```

#### c. (4 points) Analyze the seasonality of storm intensity. Create a box plot displaying the distribution of wind speed (y-axis) for each month (x-axis). You must ensure month is treated as a factor (categorical variable) so the boxes appear correctly. Fill the boxes with a color of your choice (e.g., "lightblue"). Add a title "Wind Speed Distribution by Month" and appropriate axis labels.

```{r}
storms <- mutate(storms, month = as.factor(month))
p <- storms |> 
  ggplot(
  aes(x = month, y = wind)
  ) + 
  geom_boxplot(fill = "lightblue") +
  labs(
    title = "Wind Speed Distribution by Month",
    x = "Month", y = "Wind speed (knots)"
    )
print(p)
```

#### d. (4 points) Visualize the relationship between pressure and wind. Create a scatter plot where points are colored by the `storm_category` you created in part (a). Add a smooth trend line (linear regression) for each category to see if the pressure-wind relationship differs by storm intensity.

```{r}
p <- storms |> 
  ggplot(
  aes(x = pressure, y = wind)
  ) + 
  geom_point(aes(col = storm_category), size = 0.1) +
  geom_smooth(aes(col = storm_category), method = 'lm', se = FALSE)
print(p)
```

### 5. (15 points) Utilize the `starwars` dataset in `dplyr` package to answer the following sub-questions.

```{r}

# Load the required packages and dataset
library(tidyverse)
data("starwars")

```

#### a. (3 points) Use the `ggplot2` package to create a histogram of characters' height. Add a vertical line representing the median height of the entire dataset.

```{r}
# Histogram for character's height (missing data dropped)
p <- starwars |> filter(!is.na(height)) |>
  ggplot(aes(height)) +
  geom_histogram(binwidth = 10, color = "grey", fill = "lightblue") +
  geom_vline(aes(xintercept = median(height)), lty = "dashed")
print(p)

```

#### b. (4 points) Use the `ggplot2` package to create a scatter plot showing the relationship between height (x-axis) and mass (y-axis) with a label for the outlier in the plot. Who is the outlier? Explain in one sentence why this outlier makes the rest of the data hard to see.

```{r}
p <- starwars |> filter(!is.na(height) & !is.na(mass)) |>
  ggplot(aes(x = height, y = mass)) +
  geom_point() +
  geom_text(
    aes(label= ifelse(mass >= quantile(mass, 0.99), "Outlier", "")),
    hjust=-0.1, vjust=0.5, col = "red"
  )
print(p)

# This outlier extended the y-axis, making other data crowded in one small vertical range
```

#### c. (4 points) Improve the visibility of the scatter plot in part (b) by applying Visualization Principle. (Hints: Filter the data to exclude characters with a mass greater than 500kg and map `gender` to the color aesthetic. Then add a trend line for extra insight.)

```{r}
p <- starwars |> filter(!is.na(height) & !is.na(mass) & !is.na(gender)) |>
  filter(mass <= 500) |> 
  ggplot(aes(x = height, y = mass)) +
  geom_point(aes(col = gender)) +
  geom_smooth(aes(col = gender), method = 'lm', se = FALSE) +
  ggtitle("Filtered mass > 500kg (n=1)")
print(p)

```

#### d. (4 points) Use `dplyr` to create a ranked bar chart of the Top 10 Species with the highest average height (exclude species with only 1 character).

```{r}
p <- starwars |> filter(!is.na(height) & !is.na(species)) |>
  group_by(species) |> summarize(avg_height = mean(height), n_char = n()) |>  
  filter(n_char > 1) |>
  mutate(species = reorder(species, -avg_height)) |> head(10) |>
  ggplot(aes(species, avg_height)) +
  geom_bar(stat = 'identity')
print(p)

```

### 6. (15 points) Track disease outbreaks using the `us_contagious_diseases` dataset included in the `dslabs` package. This dataset includes yearly counts of diseases by state. You will analyze the spread of Measles before the vaccine was introduced.

```{r}

# Load the required packages and dataset
library(dslabs)
library(dplyr)
library(ggplot2)
data("us_contagious_diseases")

```

#### a. (4 points) Data Preparation. 1) Filter the dataset to include only the disease "Measles". 2) Create a new column called `rate_per_10000` using the formula: count/population times 10000. 3) Remove any rows where `rate_per_10000` is missing (NA). 4) Store this cleaned data in a new object called `measles_data` hence print the first 5 rows.

```{r}
measles_data <- us_contagious_diseases |> 
# 1. Filter measles
  filter(disease == "Measles") |> 
# 2. Create "rate_per_10000"
  mutate(rate_per_10000 = count/population*10000) |>
# 3. Remove missing rows
  filter(!is.na(rate_per_10000))
# 4. First 5 rows
print(head(measles_data,5))
```

#### b. (4 points) Compare the Measles infection rate trends of "California" and "New York" over time. 1) Filter your `measles_data` to include only rows where the state is "California" or "New York". 2) Create a line plot with year on the x-axis and rate_per_10000 on the y-axis. 3) Use color to distinguish between the two states. 4) Add a title "Measles Rate Comparison: CA vs NY".

```{r}
p <- measles_data |> 
  filter(state %in% c("California","New York")) |>
  ggplot(aes(x=year, y=rate_per_10000, col=state)) +
  geom_line() +
  ggtitle("Measles Rate Comparison: CA vs NY")
print(p)
```

#### c. (4 points) Find the worst outbreak year for the entire country. 1) Using `measles_data`, group by year. 2) Summarize the total number of count (cases) for the whole US in that year. Call this `total_cases`. 3) Arrange the results in descending order to find the year with the highest number of cases. 4) Print the top 1 row (the worst year).

```{r}
worst_outbreak <- measles_data |> 
  group_by(year) |> 
  summarize(total_cases=sum(count)) |>
  arrange(-total_cases) |>
  head(1)
print(worst_outbreak)
```

#### d. (3 points) Identify the most prevalent disease in the dataset. 1) Group the entire `us_contagious_diseases` dataset by disease. 2) Calculate the total number of cases (count) for each disease across all years and states. 3) Sort the results in descending order to identify which disease has the highest cumulative number of cases hence print the resulting table.

```{r}
cum_cases <- us_contagious_diseases |>
  group_by(disease) |>
  summarize(total_cases=sum(count)) |>
  arrange(-total_cases)
print(cum_cases)
```
