---
title: "COMP2501 Assignment 2"
output:
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Requirements

**Submission deadline: Mar 31st, 2026 at 23:59.**

**Full mark of assignment 2: 50.**

For the following questions, please:

1.  Replace all [Input here] places with your information or your answer.
2.  Complete the code block by adding your own code to fulfill the requirements in each question. Please use the existing code block and do not add your own code block. Noting that please use `head()` to show the corresponding results if there are too many rows in them.

Please make sure your Rmd file is a valid Markdown document and can be successfully knitted.

For assignment submission, please knit your final Rmd file into a Word document, and submit both your **Rmd** file and the knitted **Microsoft Word** document file to Moodle. You get 0 score if 1) the Rmd file you submitted cannot be knitted, and 2) you have not submitted a Word document. For each visualization question, please make sure that the generated plot is shown in-place with the question and after the code block.

------------------------------------------------------------------------

## Name and UID

Name: Kan Chi Fu

UID: 3035807902

------------------------------------------------------------------------

### Environmental setup

You need to have the `datasets`, `dslabs`, `tidyr`, `dplyr`, `rvest`, `stringr`, `lubridate`, `janeaustenr`, `tidytext`, `textdata` and `ggplot2` packages installed. If not yet, please run `install.packages(c("datasets", "dslabs", "tidyr", "dplyr", "rvest", "stringr", "lubridate", "janeaustenr", "tidytext", "textdata", "ggplot2"))` in your R environment.

```{r}
# Load the package.
library(datasets)
library(dslabs)
library(tidyr)
library(dplyr)
library(rvest)
library(stringr)
library(lubridate)
library(janeaustenr)
library(tidytext)
library(textdata)
library(ggplot2)

```

### 1. (3 points) Load the built-in `airquality` dataset and view its first 5 rows. 1) Reshape the dataset (named `airquality_long`) using the `pivot_longer` function to convert the variables `Ozone`, `Solar.R`, `Wind`, and `Temp` into a new column named `Measurement`, with corresponding values in a new column named `Value`. 2) Reshape the `airquality_long` dataset (named `airquality_unite`) using the `unite` function to combine the `Month` and `Day` columns (with `-` as a separator) into a new column named `Date`.

```{r}
data("airquality")
#head(airquality,5)
airquality_long <- airquality |>
  pivot_longer(c(Ozone,Solar.R,Wind,Temp),names_to = 'Measurement',values_to = 'Value') |>
  unite('Date',Month,Day,sep = '-')
head(airquality_long,5)
```

### 2. (6 points) Reshape data using the `billboard` dataset. The `billboard` dataset records the rank of songs in the year 2000. The data is in a `wide` format. You need to convert it to a `long` format to analyze the trends.

```{r}
data("billboard")
head(billboard,5)
```

#### a. (3 points) Convert the `billboard` dataset from `wide` to `long` format. 1). Use the `pivot_longer` function on the `billboard` dataset. 2). Select all columns from `wk1` to `wk76`. 3). Put the column names into a new column called `week`. 4). Put the values into a new column called `rank`. 5). Remove any rows where the `rank` is missing (NA). 6). Save the result as a new data frame called `long_billboard`. Print the first 5 rows.

```{r}
long_billboard <- billboard |>
  pivot_longer(`wk1`:`wk76`,names_to = 'week',values_to = 'rank',values_drop_na = T)
head(long_billboard,5)
```

#### b. (3 points) Find the longest-charting song. 1). Use the `long_billboard` dataset. 2). Group the data by `artist` and `track`. 3). Count the total number of weeks each song stayed on the chart. 4). Sort the result in descending order. 5). Print the top 1 song.

```{r}
charting_weeks <- long_billboard |>
  group_by(artist,track)|>
  summarise(number_of_weeks = n()) |>
  arrange(-number_of_weeks)
head(charting_weeks,1)
```

### 3. (6 points) Combine and analyze election data using the `polls_us_election_2016` and `results_us_election_2016` datasets in the `dslabs` package. These datasets contain polling data and actual results from the 2016 US Presidential Election. You need to summarize the polls and join them with the actual results to evaluate poll accuracy.

```{r}
data("polls_us_election_2016")
head(polls_us_election_2016)
```

#### a. (3 points) Summarize and join the data. 1). Use the `polls_us_election_2016` dataset. 2). Filter the data to keep only rows where the grade is "A" or "A-". 3). Group the data by `state`. 4). Calculate the average `rawpoll_clinton` and average `rawpoll_trump` for each state. 5). Save this as a new data frame called `state_polls`. 6). Use the `inner_join` function to combine `state_polls` with the `results_us_election_2016` dataset using the `state` column. 7). Save the joined data frame as `poll_results`. Print the first 5 rows.

```{r}
state_polls <- polls_us_election_2016 |>
  filter(grade=='A'|grade=='A-') |>
  group_by(state) |>
  summarise(
    avg_poll_clinton = mean(rawpoll_clinton),
    avg_poll_trump = mean(rawpoll_trump)
  )
head(state_polls,5)
poll_results <- state_polls |>
  inner_join(results_us_election_2016,by = 'state')
head(poll_results,5)
```

#### b. (3 points) Visualize the polling accuracy. 1). Use the `poll_results` data frame. 2). Create a scatter plot using `ggplot2`. 3). Put the actual `clinton` result on the x-axis and the `avg_poll_clinton` on the y-axis. 4). Color the points by `state`. 5). Add a smooth trend line using the linear model method to show the overall relationship. 6). Add a title "Actual Results vs Polling Predictions for Clinton" and add axis labels.

```{r}
poll_results |>
  ggplot(aes(x = clinton, y = avg_poll_clinton)) +
  geom_point(aes(col = state)) +
  geom_smooth(method = 'lm', se = FALSE) +
  ggtitle('Actual Results vs Polling Predictions for Clinton') +
  theme(
    legend.title = element_text(size = 5),
    legend.text = element_text(size = 5)
  )
```

### 4. (5 points) Scrape data from Wikipedia with the `rvest` package, and answer the following questions.

#### a. (2 points) Choose a Wikipedia page of your interest (for example, "Data Science" in <https://en.wikipedia.org/wiki/Data_science>). Scrape the first paragraph from the Introduction section of that page. Store the scraped text in a variable and print it.

```{r}
h <- read_html('https://en.wikipedia.org/wiki/Chiikawa') 
p <- h |> 
  html_elements('div') |>
  html_elements('.mw-content-ltr') |>
  html_elements('p')
p_text <- p[2] |>
  html_text()
p_text
```

#### b. (3 points) Visit the Wikipedia page of "Data Science" (<https://en.wikipedia.org/wiki/Data_science>), try to get the image URL of the image of "The existence of Comet NEOWISE". In your scarping you can have more than one URL, as long as the required URL is contained in the output. Please output the full image URL (starting with "<https://>") rather than the URL starting with "//".

```{r}
h <- read_html('https://en.wikipedia.org/wiki/Data_science') 
f <- h |> 
  html_elements('div') |>
  html_elements('.mw-content-ltr') |>
  html_elements('figure')
f_src <- f[1] |>
  html_elements('img') |>
  html_attr('src')
f_src <- paste0('https:',f_src)
f_src
```

### 5. (3 points) Use the `lubridate` package in R, parse the `date_time` column in the `date_data` and create new columns for standard `date` and `time` components, and print the final results.

```{r}
date_data <- data.frame(date_time = c("2026-02-16 7:30:15", "2023-02-17 12:15:30", "2026-02-18 23:59:59"))
date_df <- date_data |>
  mutate(
    date_time = ymd_hms(date_time)
    ) |>
  mutate(
    date = date(date_time),
    time = hms::as_hms(date_time)
    #time = format(date_time,'%H:%M:%S')
  )
date_df
```

### 6. (3 points) Use the `stringr` package in R, perform the following tasks: 1) Extract all the phone numbers from the following text: "Please call us at 123-456-7890 or 555-555-5555." 2) Extract all the email addresses from the following text: "Contact us at [info\@example.com](mailto:info@example.com){.email} or [support\@example.com](mailto:support@example.com){.email}." 3) Replace all the URLs in the following text with the string "URL": "Check out our website at <https://www.example.com> and our blog at <https://blog.example.com>.". Print the corresponding results.

```{r}
str1 <- 'Please call us at 123-456-7890 or 555-555-5555.'
str2 <- 'Contact us at info@example.com or support@example.com.'
str3 <- 'Check out our website at https://www.example.com and our blog at https://blog.example.com.'

phone_numbers <- str1 |> 
  str_extract_all('(\\d{3}-\\d{3}-\\d{4})')

email_addresses <- str2 |>
  str_extract_all('(\\w+@\\w+.com)')

replace_urls <- str3 |> 
  str_replace_all('(https://(\\w+.)+com)','URL')

phone_numbers; email_addresses; replace_urls
```

### 7. (12 points) In this question, you will use the built‑in `stringr::sentences` vector.

```{r}
sentences_vec <- stringr::sentences
sentences_tbl <- tibble(id = seq_along(sentences_vec), text = sentences_vec)

head(sentences_vec,5)
head(sentences_tbl,5)
```

#### a. (4 points) Using regex, create logical flags per sentence: 1) `has_number`: contains any digit (Arabic numerals); 2) `has_color`: contains a basic color word from the set `{red, blue, green, yellow, black, white}` (case‑insensitive); 3) `has_punct`: ends with '!' or '?'. Report counts of each flag (i.e. how many `TRUE`).

```{r}
flags <- sentences_tbl |> 
  mutate(
    has_number = str_detect(text, '\\d'),
    has_color = str_detect(text, '(red)|(blue)|(green)|(yellow)|(black)|(white)'),
    has_punct = str_detect(text, '[!|?]$')
  )
flags_count <- flags |> 
  select(has_number,has_color,has_punct) |> 
  colSums()
flags_count
```

#### b. (4 points) 1) Tokenize `sentences_tbl` into one word per row (`unnest_tokens(word, text)` from `tidytext`). 2) Remove stop words using `tidytext::stop_words`. 3) Compute the top 15 most frequent words overall (in descending order) and plot a horizontal bar chart.

```{r}
most_freq_words <- sentences_tbl |>
  unnest_tokens(word,text) |>
  filter(!word %in% tidytext::stop_words$word) |>
  count(word) |>
  arrange(desc(n)) |>
  head(15)

most_freq_words |> 
  ggplot(aes(reorder(word,n),n)) +
  geom_bar(stat = 'identity') +
  coord_flip()
```

#### c. (4 points) 1) Create bi‑grams (two‑word sequences) from the sentences. (Hint: `unnest_tokens(bigram, text, token = "ngrams", n = 2)`). 2) Remove bi‑grams where either token is a stop word. 3) Show the top 15 bigrams by count using a bar chart.

```{r}
most_freq_bigrams <- sentences_tbl |>
  unnest_tokens(bigram,text,token = 'ngrams',n = 2) |>
  separate(bigram, c('g1','g2'), sep = ' ') |>
  filter(
    !(g1 %in% tidytext::stop_words$word | g2 %in% tidytext::stop_words$word)
  ) |>
  mutate(
    bigram = paste(g1,g2)
  ) |>
  select(id,bigram) |>
  count(bigram) |>
  arrange(desc(n)) |>
  head(15)
most_freq_bigrams |> 
  ggplot(aes(reorder(bigram,n),n)) +
  geom_bar(stat = 'identity') +
  coord_flip()
```

### 8. (12 points) In this question, you will analyze a classic public‑domain text named "Pride and Prejudice" from the `janeaustenr` package.

```{r}
library(janeaustenr)

# choose one book, e.g. Pride and Prejudice
book_df <- prideprejudice %>%
  tibble(text = .) %>%
  mutate(line = row_number()) %>%
  select(line, text)
head(book_df,5)
```

#### a. (3 points) Clean the text and store it as `clean_df` by: 1) converting text to lowercase; 2) removing non‑alphabetic characters; and 3) removing blank or empty lines. Print the first 5 rows.

```{r}
clean_df <- book_df |>
  mutate(text = str_to_lower(text)) |>
  mutate(text = str_replace_all(text,'[^a-zA-Z]+',' ')) |>
  filter(str_detect(text,'\\w'))
head(clean_df,5)
```

#### b. (3 points) Tokenize the cleaned text into individual words using `unnest_tokens()`. Remove stop words using `stop_words` from `tidytext`. Store the result as `words_df`. Print the first 5 rows.

```{r}
words_df <- clean_df |>
  unnest_tokens(word,text) |>
  filter(!word %in% tidytext::stop_words$word)
head(words_df,5)
```

#### c. (3 points) Compute word frequencies. Create a tibble `top20_words` containing the 20 most frequent words. Print the result.

```{r}
top20_words <- words_df |>
  count(word) |>
  arrange(desc(n)) |>
  head(20)
top20_words
```

#### d. (3 points) Create a horizontal bar plot of the top 20 most frequent words.

```{r}
top20_words |>
  ggplot(aes(reorder(word,n),n)) +
  geom_bar(stat = 'identity') +
  coord_flip()
```
